IRS Migration Project
========================================================

There are three research questions to address in this project:
1. What aspects of an area pull people in?
  * Do people move to areas with a higher income?
  * Do people move to areas with greater racial/ethnic diveristy
  * Do people move to areas with greater education levels
  * Do people move to areas with greater employment growth
  * Do people move to areas with a lower unemployment rate
  * Do people move to areas that are more urban
  * Do people move to areas that are part of a metropolitan/megapolitan/global city area
2. What aspects of an area push people out? (same sorts of things as above)
3. Can a network be seen indicating a national hierarchy mirroring the global hierarchy of cities?
  * Given the origin-destination pairs of the data what is the national county/metro hierarchy?
  * Are there sub-national hierarchies? (Are poeple from the south moving to other specific areas in the south, ect)
4. How have these things changed over time?
  * How have things changed since the recession?
The data to be used for this project comes from the IRS migration files


What aspects of an area pull people in?
--------------

```{r setup, echo=FALSE, results='hide', cache=TRUE}
library(ggplot2)
library(lme4)
library(randomForest)
library(rpart)
attach(mig3)
```

```{r spatial, echo=FALSE, results='markup', comment=""}


# need to disagregate the data - there is too much going on, or rewrite the controls
```

```{r table, results='markup', comment="", echo=FALSE}
library(xtable)
tab1 <- full3[which (full3$year==2004), c(13, 2, 10, 11)]
tab1.2004 <- head(tab1[order(tab1$Mrank),])

tab1 <- full3[which (full3$year==2005), c(13, 2, 10, 11)]
tab1.2005 <- head(tab1[order(tab1$Mrank),])

tab1 <- full3[which (full3$year==2006), c(13, 2, 10, 11)]
tab1.2006 <- head(tab1[order(tab1$Mrank),])

tab1 <- full3[which (full3$year==2007), c(13, 2, 10, 11)]
tab1.2007 <- head(tab1[order(tab1$Mrank),])

tab1 <- full3[which (full3$year==2008), c(13, 2, 10, 11)]
tab1.2008 <- head(tab1[order(tab1$Mrank),])

tab1 <- full3[which (full3$year==2009), c(13, 2, 10, 11)]
tab1.2009 <- head(tab1[order(tab1$Mrank),])

tab <- cbind(tab1.2004, tab1.2005, tab1.2006)
tab.1 <- cbind(tab1.2007, tab1.2008, tab1.2009)

xtable(tab.1)
```



```{r hierarchical modelsOD}
mod1 <- lmer(Tmig ~ 1 + (1 | year2), data= mig3)
#mod2 <- lmer(Tmig ~ T_EMP_D + D_jobsLT29y + D_jobs30_54 + D_jobsGT55 + (1 | year2), data= mig3)
#mod3 <- lmer(Tmig ~ T_EMP_D + D_Tjobs + (1 | year2), data= mig3)
#mod4 <- lmer(Tmig ~ D_earnGT3333 + D_earnTo3333 + D_earnLT1250 + (1 | year2), data= mig3)
#mod5 <- lmer(Tmig ~ T_EMP_D + D_Tjobs + D_Pop + (1 | year2), data= mig3)
```

```{r HLMfull3}
library(lme4)
mod1 <- lmer(Tmig ~ 1 + (1 | year), data=full3)
#mod2 <- lmer(Tmig ~ AGI + Conections + T_EMP + (1 | year), data=full3)
#mod3 <- lmer(Tmig ~ empLT29 + emp30_54 + empGT55 + (1 | year), data=full3)
#mod3.1 <- lmer(Tmig ~ earnLow +earnMid + earnHigh + (1 | year), data=full3)
#mod4 <- lmer(Tmig ~ T_EMP + Wages_Sal + Dividend + Interest + (1 | year), data=full3)
#mod5 <- lmer(Tmig ~ T_EMP + AGI + Conections + (cMig | year), data=full3)

#lm1 <- lm(Tmig ~   Pop+Tmig+Mrank+Crank+Metro+Micro+AGI+Aggr_AGI+Wages_Sal+Dividend+Interest+T_EMP+earnLow+earnMid+earnHigh+Jden+Pden, data=s2004)
```


```{r plots}
library(ggplot2)
p <- ggplot(full3, aes(as.factor(year), Tmig))
p+geom_boxplot()+geom_jitter()
b <- ggplot(full3, aes(V5, Tmig))
b+geom_boxplot()+geom_jitter()

boxplot(Tmig ~ year, full3)

panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste(prefix, txt, sep="")
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~ T_EMP + Tmig + year + AGI + Mrank + Crank + Conections , data= full3, lower.panel=panel.smooth, upper.panel = panel.cor, pch=10)
```
empLT29 + emp30_54 + empGT55 + earnLow +earnMid + earnHigh


